---
title: "&#127826;CherryModels"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{CherryModels}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<hr>

The <b>CherryModels</b> package facilitates computing thousands of linear mixed-effects models to determine the model that best explains your data. To this end, all combinations of fixed effects with and without interaction are generated and computed with `lmerTest::lmer()`. Subsequently, a step-wise model reduction is performed on every model with `lmerTest::step()`.  

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=FALSE, warning=FALSE}
library(CherryModels)
```

We use the `mpg` data set provided in the `ggplot2` package for demonstration.
```{r}
glimpse(mpg)
```

This tibble is our *model_data* and will be passed to `lmer`. Besides the data, this function also needs a formula: <br> 
`response ~ fixed + random`

- The *response* is the dependent variable. Although a transformation inside the model formula is possible, e.g. `response = "log(my_var)"`, it is not recommended since it may cause problems in the `show_report()` function.

- Fixed effects (independent variables) that are to be investigated for interaction effects are passed to the argument *fixed_inter*. 

- The maximum number of these effects that should be considered for interaction is specified in *max_comb*. As an example, let `c("A", "B", "C", "D")` be the fixed effects. If *max_comb* is set to `2`, then the interaction effects generated are `A:B`, ..., `C:D`. 

- The argument *comb_comb* specifies the maximum number of interaction terms in the same manner, e.g. `comb_comb = 2` means `A:B + A:C`, ..., `B:D + C:D`.

- Fixed effects that should be included without interaction can be supplied to *fixed_no_inter*.

- The random effects are supplied to *random* (please be aware of the notation required by `lmer`).

A <i><b>dry</b></i> run can be used to display the generated formulas without computing the models.

```{r}
formulas <- cherry_models(model_data = mpg,
                          response = "cty",
                          fixed_inter = c("displ", "year", "cyl"),
                          fixed_no_inter = c("trans", "drv"),
                          max_comb = 3,
                          comb_comb = 2,
                          random = c("(1 | manufacturer/model)"),
                          n.workers = 2,
                          dry = TRUE,
                          background = FALSE)

head(formulas)
```

By default the computation is run in parallel with two threads and can be specified with the *n.workers* argument. If the number of cores of the processor is unknown the function `parallel::detectCores()` can be run (you may need to install the **parallel** package).

Let's run the model:

```{r, warning=FALSE, message=FALSE}
x <- cherry_models(model_data = mpg,
                   response = "cty",
                   fixed_inter = c("displ", "year", "cyl"),
                   fixed_no_inter = c("trans", "drv"),
                   max_comb = 3,
                   comb_comb = 2,
                   random = c("(1 | manufacturer/model)"),
                   n.workers = 2,
                   dry = FALSE,
                   background = FALSE)
```

The call will print the number of formulas it is computing and subsequently the process of the computation.

The returned data.frame has one row per model containing the extracted values of the initial (before `step()`) and the final model (after `step()`).

```{r, warning=FALSE, message=FALSE}
glimpse(x)
```

*Reduced* shows the ratio of effects dropped during model reduction to initial number of fixed effects included. *Fixed* shows the fixed effects retained in the final model.

Additionally, metadata are saved in the attributes. These include the values passed to the arguments.

```{r, warning=FALSE, message=FALSE}
attributes(x)
```

<hr>

Despite parallelization, computation can take some time. To avoid having your session being busy for hours, you can run the computation as a background job (see package `callr` for details) by setting *background* to `TRUE`. In this case the function returns a so-called *handler*, i.e. a connection to another R session running the computation in the background. To communicate with this session you can use `get_process_result()` on the object. This will print the latest messages output by the background process. Once the the process is finished, this function returns the result of the computation.

```{r, warning=FALSE, message=FALSE}
#> I don't want this code chunk to be run during the generation of this document, hence the 
#> if (FALSE) {...} statement, but you are safe to copy-paste it into your console and run it.
if (FALSE) { 

  proc <- cherry_models(model_data = mpg,
                        response = "cty",
                        fixed_inter = c("displ", "year", "cyl"),
                        fixed_no_inter = c("trans", "drv"),
                        max_comb = 3,
                        comb_comb = 2,
                        random = c("(1 | manufacturer/model)"),
                        n.workers = 2,
                        dry = FALSE,
                        background = TRUE)
  
  x <- get_process_results(proc)

}
```

When starting the call:

```{r}
#> Starting r_process in background and returning handler
```

Calling the object (here `proc`) shows some more information:

```{r}
#> proc
#> PROCESS 'R', running, pid 91556.
```

When the computation is finished:

```{r}
#> proc
#> PROCESS 'R', finished.
```

While the process is not finished, we get information on what is happening by calling `get_process_result()`.

First it is loading packages...

```{r}
#> x <- get_process_results(proc)
#> busy:
#> Loading required package: Matrix
```

...then the model formulas are generated...

```{r}
#> x <- get_process_results(proc)
#> busy:
#> generating 63 models
```

...subsequently the models are computed and the progress is reported...

```{r}
#> x <- get_process_results(proc)
#> busy:
#>  Progress: ───────────────────────────                                      100%
```

...and finally the process is finished. The object `x` now contains the data.frame containing the results and the object we passed to the function, i.e. the handler, is removed from the global environment.

```{r}
#> x <- get_process_results(proc)
#> finished
#> removing proc
```

<hr>

To inspect the results we can call `show_report()` which will start a *shiny* app. In the app click the button with the <b>i</b> icon in the top right corner for further explanation.

```{r}
if (FALSE) show_report(x)
```

<hr>
<br>
<br>
<br>
